---
title: "UW Grade Distributions: INFO vs CSE (UW 2016)"
author: "Khadija Saho"
date: 01/25/205
format:
  html:
   embed-resources: true
---

## Library
```{r}
#| warning: false
#| include: false
#I will be using dplyr 

library("dplyr")

```
##Code to load grade distribution data files:

###Data originally from reddit user austingwalters in this reddit thread: https://www.reddit.com/r/udub/comments/4e6kxl/i_made_a_website_that_shows_grade_distributions/
# Dataset and 2016 FOIA request info from their dropbox: https://www.dropbox.com/scl/fo/8lbyazjlt5jmv86jvvs06/AGtgMaWeyZtZuS2fSYLYn_4?dl=0&e=1&rlkey=1sg7t372nhslxogwjl3e9lt9t

```{r}

cse_grades_df <- read.csv("https://raw.githubusercontent.com/kylethayer/Info-201-au-24-datasets/refs/heads/main/CSE_grades_2016.csv")

info_grades_df <- read.csv("https://raw.githubusercontent.com/kylethayer/Info-201-au-24-datasets/refs/heads/main/INFO_grades_2016.csv")
```


# Rows 
```{r}
glimpse(info_grades_df)
head(info_grades_df, 10)

glimpse(cse_grades_df)
head(cse_grades_df, 10)
```


```{r}
# Look at column names
names(info_grades_df)
```

# The structure of the data

```{r}
str(info_grades_df)
```

#  The first few rows

```{r}
head(info_grades_df)
```


# INFO Answers for (a-f)

```{r}
nrow(info_grades_df) #a
mean(info_grades_df$Student_Count, na.rm = TRUE) #b
sum(info_grades_df$Student_Count, na.rm = TRUE) #c
min(info_grades_df$Average_GPA, na.rm = TRUE) #d
max(info_grades_df$Average_GPA, na.rm = TRUE) #e
mean(info_grades_df$Average_GPA, na.rm = TRUE)#f

```

# CSE Answers for (a-f)
```{r}
nrow(cse_grades_df) #a
mean(cse_grades_df$Student_Count, na.rm = TRUE) #b
sum(cse_grades_df$Student_Count, na.rm = TRUE) #c
min(cse_grades_df$Average_GPA, na.rm = TRUE) #d
max(cse_grades_df$Average_GPA, na.rm = TRUE) #e
mean(cse_grades_df$Average_GPA, na.rm = TRUE)#f

```

# INFO summary 
```{r}
info_total_classes  <- nrow(info_grades_df)
info_avg_students   <- mean(info_grades_df$Student_Count, na.rm = TRUE)
info_total_students <- sum(info_grades_df$Student_Count, na.rm = TRUE)
info_min_gpa        <- min(info_grades_df$Average_GPA, na.rm = TRUE)
info_max_gpa        <- max(info_grades_df$Average_GPA, na.rm = TRUE)
info_mean_gpa       <- mean(info_grades_df$Average_GPA, na.rm = TRUE)

info_summary <- data.frame(
  Metric = c(
    "Total class records",
    "Average students per class",
    "Total student count (summed)",
    "Minimum average GPA",
    "Maximum average GPA",
    "Mean average GPA"
  ),
  Value = round(c(
    info_total_classes,
    info_avg_students,
    info_total_students,
    info_min_gpa,
    info_max_gpa,
    info_mean_gpa
  ), 2)
)

info_summary
```
# CSE summary
```{r}

cse_total_classes  <- nrow(cse_grades_df)
cse_avg_students   <- mean(cse_grades_df$Student_Count, na.rm = TRUE)
cse_total_students <- sum(cse_grades_df$Student_Count, na.rm = TRUE)
cse_min_gpa        <- min(cse_grades_df$Average_GPA, na.rm = TRUE)
cse_max_gpa        <- max(cse_grades_df$Average_GPA, na.rm = TRUE)
cse_mean_gpa       <- mean(cse_grades_df$Average_GPA, na.rm = TRUE)

cse_summary <- data.frame(
  Metric = c(
    "Total class records",
    "Average students per class",
    "Total student count (summed)",
    "Minimum average GPA",
    "Maximum average GPA",
    "Mean average GPA"
  ),
  Value = round(c(
    cse_total_classes,
    cse_avg_students,
    cse_total_students,
    cse_min_gpa,
    cse_max_gpa,
    cse_mean_gpa
  ), 2)
)

cse_summary

```

# Program Summaries

## Informatics (INFO)

The Informatics (INFO) program has `r round(info_total_classes, 2)` total class records in the 2016 dataset. Each class record represents a specific course, which means the same course can appear multiple times across different quarters or sections. The average Student_Count per INFO class record is `r round(info_avg_students, 2)` students. The total sum of Student_Count is `r round(info_total_students, 2)`, this shows the overall enrollment activity in INFO courses but it is likely counts individual students twice ,  a single student may be enrolled in multiple INFO classes or in the same course offered in different quarters.

For grading, the minimum Average_GPA for INFO class records is `r round(info_min_gpa, 2)`, which may show that the clas may be challenging or strictly graded course. The maximum Average_GPA is `r round(info_max_gpa, 2)`, which could be because the class is smaller class, a more lecture type course, or a course that is more lenient in grading.The overall mean Average_GPA across all INFO class records is `r round(info_mean_gpa, 2)`, this shows a general idea of how grades are distributed at the course level within the Informatics program.

## Computer Science & Engineering (CSE)

The Computer Science & Engineering (CSE) program has `r round(cse_total_classes, 2)` total class records in the 2016 dataset, which shows individual course offerings across quarters and sections. The average Student_Count per CSE class record is `r round(cse_avg_students, 2)` students, whihc shows the typical class sizes within the program. The total sum of Student_Count across all CSE class records is `r round(cse_total_students, 2)`, which shows the overall scale of enrollment but it is likely counts individual students twice who are enrolled in multiple CSE courses.

For the grading, the minimum Average_GPA for CSE class records is `r round(cse_min_gpa, 2)`, show that course maybe  difficult or strictly graded courses. The maximum Average_GPA is `r round(cse_max_gpa, 2)`, which may correspond to smaller or more specialized offerings. The mean Average_GPA across all CSE class records is `r round(cse_mean_gpa, 2)`, show the overall grading outcomes across CSE courses at the course level rather than the individual student level.

# Compare Summary 
Based on the 2016 data, the INFO program has `r info_total_classes` class records, while the CSE program has `r cse_total_classes` class records, which means that CSE has a larger number of recorded offerings in this dataset. The average class size is `r round(info_avg_students,2)` students for INFO and `r round(cse_avg_students,2)` students for CSE, showing the differences in enrollment scale between the two programs. When it comes to grading, INFO has a mean Average_GPA of `r round(info_mean_gpa,2)` and CSE has a mean Average_GPA of `r round(cse_mean_gpa,2)`. This means that, on average, grades in INFO courses tend to be little higher than those in CSE courses. These values reflect course level averages and which may be influenced by course type, level, and grading practices.

# Questions
How is the Average_GPA values different between lower-division and upper-division courses within each program?

Is there a relationship between class size (Student_Count) and Average_GPA, could it be that smaller classes tend to have higher average GPAs than larger classes?
